---
title: "Campbell_2017_seurat_analysis"
author: "Matthew Banton"
date: "15/11/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load packages and set working directory

```{r library}

library(tidyverse)
library(scater)
library(dplyr)
library(limma)
library(SingleCellExperiment)
library(Seurat)
library(mclust)
library(Matrix)
library(tibble)
library(PerformanceAnalytics)

## set working directory
setwd("/media/data/mattb/projects/Brian_scRNAseq_website/R_projects/mouse_hypothalamus_scRNAseq")

```

# Analysis of Campbell scRNAseq data for Chow fed mice (5 replicates) using seurat


### Load CHOW mouse data

```{r readRDS}

## Load seurat object
seurat_campbell_chow<- readRDS(file = "./seurat_campbell_chow_just_created.rds")

seurat_campbell_chow


```

### Quality control on CHOW mouse data

```{r QC}

## Import list of mouse mitocondrially encoded genes:
mito_genes<- read.csv("mito_genes_table.csv", header = TRUE)


## Make list of mitocondrial gene IDS:
mito_genes <- mito_genes$id_with_url


## Identify which mitochondrially expressed gene IDs are present in my dataset:
mito_genes_present<-seurat_campbell_chow@raw.data[mito_genes, ]

mito_genes_present<- mito_genes[!grepl("NA", rownames(mito_genes_present))]

seurat_campbell_chow@raw.data[mito_genes_present, 1:5]

dim(seurat_campbell_chow@raw.data[mito_genes_present, ])


## Calculate the percentage of mitcondrial gene counts per cell
percent_mito <- Matrix::colSums(seurat_campbell_chow@raw.data[mito_genes_present, ])/Matrix::colSums(seurat_campbell_chow@raw.data)


## Basic stats of proportion of mitocondrial gene expression per cell
summary(percent_mito)


## add percentage mitocondrial genes into metadata
seurat_campbell_chow <- AddMetaData(object = seurat_campbell_chow,
                    metadata = percent_mito,
                    col.name = "percent_mito")


## Look at the seurat object meta data
head(seurat_campbell_chow@meta.data)


## QC plots of number of genes, UMIs, and % mitochondria
VlnPlot(object = seurat_campbell_chow,
        features.plot = c("nGene", "nUMI", "percent_mito"),
        nCol = 3,
        x.lab.rot = TRUE,
        point.size.use = 0.2
        )


## QC plots to show the relationship between nUMIs and relative mitocondrial gene expression or number of genes.
par(mfrow = c(1, 2))
# GenePlot(object = seurat_campbell_chow, gene1 = "nUMI", gene2 = "percent_mito", do.hover = TRUE, pch.use = 1)
# GenePlot(object = seurat_campbell_chow, gene1 = "nUMI", gene2 = "nGene", do.hover = TRUE, pch.use = 1)
GenePlot(object = seurat_campbell_chow, gene1 = "nUMI", gene2 = "percent_mito", pch.use = 16, cex.use = 0.5)

GenePlot(object = seurat_campbell_chow, gene1 = "nUMI", gene2 = "nGene", pch.use = 16, cex.use = 0.5)


```

### Filter cells after QC

Filter cells out with more than 0.4% of total gene expression comming from mitocondrially encoded genes and more than 6000 genes expressed. 

```{r FilterCells}

## manual check; I already know all cells have >800 genes
table(seurat_campbell_chow@meta.data$percent_mito < 0.004 & seurat_campbell_chow@meta.data$nGene<6000)
 

# FALSE  TRUE 
#  358 10897 
 
## Filter cells with <0.4% percent_mito and <6000 genes
seurat_campbell_chow <- FilterCells(object = seurat_campbell_chow,
                    subset.names = c("nGene", "percent_mito"),
                    low.thresholds = c(800, -Inf),
                    high.thresholds = c(6000, 0.004))
 
seurat_campbell_chow
# An object of class seurat in project CAMPBELL_CHOW 
# 30000 genes across 10897 samples.
# 358 cells are filtered out; numbers consistent with above
 
```


### Log normalise gene expression per cell

```{r NormalizeData}

## Plot graph of total expression before normalisation
hist(colSums(seurat_campbell_chow@data),
     breaks = 100,
     main = "Total expression before normalisation",
     xlab = "Sum of expression")

## Normalise gene expression per cell
seurat_campbell_chow <- NormalizeData(object = seurat_campbell_chow, normalization.method = "LogNormalize", 
    scale.factor = 1e4)

## Plot graph of total expression after normalisation
hist(colSums(seurat_campbell_chow@data),
     breaks = 100,
     main = "Total expression after normalisation",
     xlab = "Sum of expression")


```



### Find genes whose expression varies between cells.

Find genes whose expression varies between cells, which will be used to construct principal componets between cells that will be used for clustering.

```{r FindVariableGenes}

## Find variable genes by expression
seurat_campbell_chow <- FindVariableGenes(object = seurat_campbell_chow,
                          mean.function = ExpMean,
                          dispersion.function = LogVMR,
                          x.low.cutoff = 0.05,
                          x.high.cutoff = 4,
                          y.cutoff = 0.75,
                          num.bin = 20,
                          binning.method = "equal_width"
                          )


# number of variable genes
length(seurat_campbell_chow@var.genes)

```

seurat_campbell_chow <- FindVariableGenes(object = seurat_campbell_chow,
                          mean.function = ExpMean,
                          dispersion.function = LogVMR,
                          x.low.cutoff = 0.05,
                          x.high.cutoff = 4,
                          y.cutoff = 0.75,
                          num.bin = 20,
                          binning.method = "equal_width"
                          )
                          
                          == 1817 variable genes

seurat_campbell_chow <- FindVariableGenes(object = seurat_campbell_chow,
                          mean.function = ExpMean,
                          dispersion.function = LogVMR,
                          x.low.cutoff = 0.02,
                          x.high.cutoff = 4,
                          y.cutoff = 0.75,
                          num.bin = 20,
                          binning.method = "equal_width"
                          )
                          
                          == 2346 variable genes

seurat_campbell_chow <- FindVariableGenes(object = seurat_campbell_chow,
                          mean.function = ExpMean,
                          dispersion.function = LogVMR,
                          x.low.cutoff = 0.02,
                          x.high.cutoff = 4,
                          y.cutoff = 0.5,
                          num.bin = 20,
                          binning.method = "equal_width"
                          )
                          
                          == 4084 variable genes


## Scale normalised gene expression per cell to remove unwanted sources of variation

Scale gene expression per cell by building linear models for nUMI, percent_mito and mouse replicate.

```{r ScaleData}

## Scale data nUMI, percent_mito and mouse replicate
seurat_campbell_chow <- ScaleData(object = seurat_campbell_chow, vars.to.regress = c("nUMI", "percent_mito", "replicate_name"))


```


### Principal component anlysis of variable genes.

Principal component anlysis of variable genes for use in cell clustering.

```{r RunPCA}

## Perform principal component analysis on variable genes
seurat_campbell_chow <- RunPCA(object = seurat_campbell_chow,
               pc.genes = seurat_campbell_chow@var.genes,
               do.print = TRUE,
               pcs.print = 1:5,
               genes.print = 5)


## visualise top genes associated with principal components
VizPCA(object = seurat_campbell_chow, pcs.use = 1:9)


## Plot principal component 1 v's 2
PCAPlot(object = seurat_campbell_chow, dim.1 = 1, dim.2 = 2)


## Plot principal component 2 v's 3
PCAPlot(object = seurat_campbell_chow, dim.1 = 2, dim.2 = 3)


## Plot heat map for gene expression of principal component 1 genes
PCHeatmap(object = seurat_campbell_chow, pc.use = 1, cells.use = 500, do.balanced = TRUE, label.columns = FALSE)
# Ignore warnings!


## Plot heat map for gene expression of principal component 2 genes
PCHeatmap(object = seurat_campbell_chow, pc.use = 2, cells.use = 500, do.balanced = TRUE, label.columns = FALSE)
# Ignore warnings!


## Plot heat map for gene expression of principal component 3 genes
PCHeatmap(object = seurat_campbell_chow, pc.use = 3, cells.use = 500, do.balanced = TRUE, label.columns = FALSE)
# Ignore warnings!


## Plot heat map for gene expression of principal component 4 genes
PCHeatmap(object = seurat_campbell_chow, pc.use = 4, cells.use = 500, do.balanced = TRUE, label.columns = FALSE)
# Ignore warnings!


## Plot multiple heatmaps of gene expression per PC genes
PCHeatmap(object = seurat_campbell_chow,
          pc.use = 5:18,
          cells.use = 500,
          do.balanced = TRUE,
          label.columns = FALSE)


## Takes a pre-computed PCA (from most variable genes identified earlier) and projects this onto the entire dataset (all genes)
seurat_campbell_chow <- ProjectPCA(object = seurat_campbell_chow, do.print = TRUE)

```



### Determine statistically significant principal components

```{r}

## Perform jackstraw statistical test to investigate statistically significant PC.
seurat_campbell_chow <- JackStraw(object = seurat_campbell_chow,
                  num.replicate = 100,
                  display.progress = TRUE
                  )
# Maximum number of PCs allowed = 20.


## Visualise JackStraw plots
JackStrawPlot(object = seurat_campbell_chow, PCs = 1:20)


## A less computationally intensive heuristic method for finding the statistically significant PCAs is using an elbow plot 
PCElbowPlot(object = seurat_campbell_chow)


```

There is little difference in the Jackstraw and elbow plot when using 1817, 2346 or 4084 variable genes.

### Cell clustering

```{r}

## Cluster cells by PC
seurat_campbell_chow <- FindClusters(object = seurat_campbell_chow, reduction.type = "pca", dims.use = 1:20, 
    resolution = 0.6, print.output = 0, save.SNN = TRUE, force.recalc = TRUE)

PrintFindClustersParams(object = seurat_campbell_chow)

seurat_campbell_chow <- RunTSNE(object = seurat_campbell_chow, dims.use = 1:20, do.fast = TRUE)

# TSNEPlot(object = seurat_campbell_chow)
TSNEPlot(object = seurat_campbell_chow,
         no.legend = TRUE,
         do.label = TRUE)
## 4084 variable genes = 19 clusters
## 2346 variable genes = 20 clusters
## 1817 variable genes = 20 clusters

table(seurat_campbell_chow@ident)

proportion_table<- table(seurat_campbell_chow@ident, seurat_campbell_chow@meta.data$replicate_name)

proportion_table<- round(prop.table(proportion_table, 2), 5)

proportion_table

TSNEPlot(object = seurat_campbell_chow,
         group.by = "replicate_name",
         no.legend = FALSE,
         do.label = FALSE,
         pt.size = 0.4
         )

proportion_table<- data.frame(matrix(proportion_table, ncol = 5))

colnames(proportion_table)<- colnames(table(seurat_campbell_chow@ident,
                                            seurat_campbell_chow@meta.data$replicate_name))

chart.Correlation(data.frame( proportion_table[1:5,]) )

#######
seurat_campbell_chow <- FindClusters(object = seurat_campbell_chow, reduction.type = "pca", dims.use = 1:20, 
    resolution = 1.0, print.output = 0, save.SNN = TRUE, force.recalc = TRUE)

PrintFindClustersParams(object = seurat_campbell_chow)

seurat_campbell_chow <- RunTSNE(object = seurat_campbell_chow, dims.use = 1:20, do.fast = TRUE)

# TSNEPlot(object = seurat_campbell_chow)
TSNEPlot(object = seurat_campbell_chow,
         no.legend = TRUE,
         do.label = TRUE)

##  4084 variable genes = 21 clusters
##  2346 variable genes = 22 clusters
table(seurat_campbell_chow@ident)

proportion_table<- table(seurat_campbell_chow@ident, seurat_campbell_chow@meta.data$replicate_name)

proportion_table<- round(prop.table(proportion_table, 2), 5)

proportion_table

TSNEPlot(object = seurat_campbell_chow,
         group.by = "replicate_name",
         no.legend = FALSE,
         do.label = FALSE,
         pt.size = 0.4
         )


proportion_table<- data.frame(matrix(proportion_table, ncol = 5))

colnames(proportion_table)<- colnames(table(seurat_campbell_chow@ident, 
                                            seurat_campbell_chow@meta.data$replicate_name))

chart.Correlation(data.frame( proportion_table[1:5,]) )


#######
seurat_campbell_chow <- FindClusters(object = seurat_campbell_chow, reduction.type = "pca", dims.use = 1:20, 
    resolution = 1.5, print.output = 0, save.SNN = TRUE, force.recalc = TRUE)

PrintFindClustersParams(object = seurat_campbell_chow)

seurat_campbell_chow <- RunTSNE(object = seurat_campbell_chow, dims.use = 1:20, do.fast = TRUE)

# TSNEPlot(object = seurat_campbell_chow)
TSNEPlot(object = seurat_campbell_chow,
         no.legend = TRUE,
         do.label = TRUE)

##  4084 variable genes = 25 clusters
##  2346 variable genes = 24 clusters

table(seurat_campbell_chow@ident)

proportion_table<- table(seurat_campbell_chow@ident, seurat_campbell_chow@meta.data$replicate_name)

proportion_table<- round(prop.table(proportion_table, 2), 5)

proportion_table

TSNEPlot(object = seurat_campbell_chow,
         group.by = "replicate_name",
         no.legend = FALSE,
         do.label = FALSE,
         pt.size = 0.4
         )


proportion_table<- data.frame(matrix(proportion_table, ncol = 5))

colnames(proportion_table)<- colnames(table(seurat_campbell_chow@ident, 
                                            seurat_campbell_chow@meta.data$replicate_name))

chart.Correlation(data.frame( proportion_table[1:5,]) )

#######
seurat_campbell_chow <- FindClusters(object = seurat_campbell_chow, reduction.type = "pca", dims.use = 1:20, 
    resolution = 2.0, print.output = 0, save.SNN = TRUE, force.recalc = TRUE)

PrintFindClustersParams(object = seurat_campbell_chow)

seurat_campbell_chow <- RunTSNE(object = seurat_campbell_chow, dims.use = 1:20, do.fast = TRUE)

# TSNEPlot(object = seurat_campbell_chow)
TSNEPlot(object = seurat_campbell_chow,
         no.legend = TRUE,
         do.label = TRUE)

##  4084 variable genes = 29 clusters
## 2346 variable genes = 28 clusters

table(seurat_campbell_chow@ident)

proportion_table<- table(seurat_campbell_chow@ident, seurat_campbell_chow@meta.data$replicate_name)

proportion_table<- round(prop.table(proportion_table, 2), 5)

proportion_table

TSNEPlot(object = seurat_campbell_chow,
         group.by = "replicate_name",
         no.legend = FALSE,
         do.label = FALSE,
         pt.size = 0.4 )

# install.packages("PerformanceAnalytics")

proportion_table<- data.frame(matrix(proportion_table, ncol = 5))

colnames(proportion_table)<- colnames(table(seurat_campbell_chow@ident, 
                                            seurat_campbell_chow@meta.data$replicate_name))

chart.Correlation(data.frame( proportion_table[1:5,]) )

# cor(proportion_table[1],proportion_table[4])

## Cluster cells using final parameters (1817 genes, 20 PC, resolution = 0.6)
seurat_campbell_chow <- FindClusters(object = seurat_campbell_chow, reduction.type = "pca", dims.use = 1:20, 
    resolution = 0.6, print.output = 0, save.SNN = TRUE, force.recalc = TRUE)


## Produce t-SNE
seurat_campbell_chow <- RunTSNE(object = seurat_campbell_chow, dims.use = 1:20, do.fast = TRUE)

```
Use 1817 genes, 20 principal components and a resolution of 0.6 to give 20 individual clusters. Less clusters are more likely to give more meaningfull functionally distinct groups of cells.


### Finding differentially expressed genes between cell clusters (cluster biomarkers)

```{r}


## Find markers for every cluster compared to all remaining cells, report both positive and negative genes.
seurat_campbell_chow_biomarkers <- FindAllMarkers(object = seurat_campbell_chow, only.pos = FALSE, min.pct = 0.2)


## Get the top 10 biomarkers per cluster
top10_seurat_campbell_markers<- seurat_campbell_chow_biomarkers %>% group_by(cluster) %>% top_n(10, avg_logFC)
top10_seurat_campbell_markers

# write.csv(as.data.frame(seurat_campbell_chow_biomarkers), file = "seurat_campbell_chow_biomarkers.csv", quote = FALSE)
write.csv(as.data.frame(top10_seurat_campbell_markers), file = "top10_seurat_campbell_chow_biomarkers.csv", quote = FALSE)




## Perform ROC DE test. This can take a long time.
seurat_campbell_chow_biomarkers_ROC <- FindAllMarkers(object = seurat_campbell_chow, only.pos = FALSE, min.pct = 0.2, test.use = "roc")

top10_seurat_campbell_markers_ROC<- seurat_campbell_chow_biomarkers_ROC %>% group_by(cluster) %>% top_n(10, avg_logFC)
top10_seurat_campbell_markers_ROC

# write.csv(as.data.frame(seurat_campbell_chow_biomarkers_ROC), file = "seurat_campbell_chow_biomarkers_ROC.csv", quote = FALSE)
write.csv(as.data.frame(top10_seurat_campbell_markers_ROC), file = "top10_seurat_campbell_chow_biomarkers_ROC.csv", quote = FALSE)

## Plot heatmap of top 10 DE genes
DoHeatmap(object = seurat_campbell_chow,
          genes.use = top10_seurat_campbell_markers$gene,
          slim.col.label = TRUE,
          remove.key = TRUE)

```



### Save seurat object of chow mice.

```{r}

## save seurat object as .rds 
#saveRDS(seurat_campbell_chow, file = "./seurat_campbell_chow_final.rds")

```















# Analysis of Campbell scRNAseq data for fasted mice (3 replicates) using seurat


### Load FASTED mouse data

```{r readRDS}

## Load seurat object
seurat_campbell_fasted<- readRDS(file = "./seurat_campbell_fasted_just_created.rds")

seurat_campbell_fasted


```

### Quality control on CHOW mouse data

```{r QC}

## Import list of mouse mitocondrially encoded genes:
mito_genes<- read.csv("mito_genes_table.csv", header = TRUE)


## Make list of mitocondrial gene IDS:
mito_genes <- mito_genes$id_with_url


## Identify which mitochondrially expressed gene IDs are present in my dataset:
mito_genes_present<-seurat_campbell_fasted@raw.data[mito_genes, ]

mito_genes_present<- mito_genes[!grepl("NA", rownames(mito_genes_present))]

seurat_campbell_fasted@raw.data[mito_genes_present, 1:5]

dim(seurat_campbell_fasted@raw.data[mito_genes_present, ])


## Calculate the percentage of mitcondrial gene counts per cell
percent_mito <- Matrix::colSums(seurat_campbell_fasted@raw.data[mito_genes_present, ])/Matrix::colSums(seurat_campbell_fasted@raw.data)


## Basic stats of proportion of mitocondrial gene expression per cell
summary(percent_mito)


## add percentage mitocondrial genes into metadata
seurat_campbell_fasted <- AddMetaData(object = seurat_campbell_fasted,
                    metadata = percent_mito,
                    col.name = "percent_mito")


## Look at the seurat object meta data
head(seurat_campbell_fasted@meta.data)


## QC plots of number of genes, UMIs, and % mitochondria
VlnPlot(object = seurat_campbell_fasted,
        features.plot = c("nGene", "nUMI", "percent_mito"),
        nCol = 3,
        x.lab.rot = TRUE,
        point.size.use = 0.2
        )


## QC plots to show the relationship between nUMIs and relative mitocondrial gene expression or number of genes.
par(mfrow = c(1, 2))
# GenePlot(object = seurat_campbell_fasted, gene1 = "nUMI", gene2 = "percent_mito", do.hover = TRUE, pch.use = 1)
# GenePlot(object = seurat_campbell_fasted, gene1 = "nUMI", gene2 = "nGene", do.hover = TRUE, pch.use = 1)
GenePlot(object = seurat_campbell_fasted, gene1 = "nUMI", gene2 = "percent_mito", pch.use = 16, cex.use = 0.5)

GenePlot(object = seurat_campbell_fasted, gene1 = "nUMI", gene2 = "nGene", pch.use = 16, cex.use = 0.5)


```

### Filter cells after QC

Filter cells out with more than 0.5% of total gene expression comming from mitocondrially encoded genes and more than 4000 genes expressed. 

```{r FilterCells}

## manual check; I already know all cells have >800 genes
table(seurat_campbell_fasted@meta.data$percent_mito < 0.005 & seurat_campbell_fasted@meta.data$nGene<4000)
 

# FALSE  TRUE 
#  143  3640 
 
## Filter cells with <0.5% percent_mito and <4000 genes
seurat_campbell_fasted <- FilterCells(object = seurat_campbell_fasted,
                    subset.names = c("nGene", "percent_mito"),
                    low.thresholds = c(800, -Inf),
                    high.thresholds = c(4000, 0.005))
 
seurat_campbell_fasted
# An object of class seurat in project CAMPBELL_FASTED 
# 21789 genes across 3640 samples.

 
```


### Log normalise gene expression per cell

```{r NormalizeData}

## Plot graph of total expression before normalisation
hist(colSums(seurat_campbell_fasted@data),
     breaks = 100,
     main = "Total expression before normalisation",
     xlab = "Sum of expression")

## Normalise gene expression per cell
seurat_campbell_fasted <- NormalizeData(object = seurat_campbell_fasted, normalization.method = "LogNormalize", scale.factor = 10000)

## Plot graph of total expression after normalisation
hist(colSums(as.data.frame(as.matrix(seurat_campbell_fasted@data))),
     breaks = 100,
     main = "Total expression after normalisation",
     xlab = "Sum of expression")


```



### Find genes whose expression varies between cells.

Find genes whose expression varies between cells, which will be used to construct principal componets between cells that will be used for clustering.

```{r FindVariableGenes}

## Find variable genes by expression
seurat_campbell_fasted <- FindVariableGenes(object = seurat_campbell_fasted,
                          mean.function = ExpMean,
                          dispersion.function = LogVMR,
                          x.low.cutoff = 0.05,
                          x.high.cutoff = 3,
                          y.cutoff = 0.75,
                          num.bin = 20,
                          binning.method = "equal_width"
                          )

# number of variable genes
length(seurat_campbell_fasted@var.genes)

```

seurat_campbell_fasted <- FindVariableGenes(object = seurat_campbell_fasted,
                          mean.function = ExpMean,
                          dispersion.function = LogVMR,
                          x.low.cutoff = 0.05,
                          x.high.cutoff = 4,
                          y.cutoff = 0.75,
                          num.bin = 20,
                          binning.method = "equal_width"
                          )
                          
                          == 1384 variable genes


seurat_campbell_fasted <- FindVariableGenes(object = seurat_campbell_fasted,
                          mean.function = ExpMean,
                          dispersion.function = LogVMR,
                          x.low.cutoff = 0.05,
                          x.high.cutoff = 4,
                          y.cutoff = 0.5,
                          num.bin = 20,
                          binning.method = "equal_width"
                          )
                          
                          == 2356 variable genes

seurat_campbell_fasted <- FindVariableGenes(object = seurat_campbell_fasted,
                          mean.function = ExpMean,
                          dispersion.function = LogVMR,
                          x.low.cutoff = 0.02,
                          x.high.cutoff = 4,
                          y.cutoff = 0.75,
                          num.bin = 20,
                          binning.method = "equal_width"
                          )
                          
                          == 1848 variable genes

seurat_campbell_fasted <- FindVariableGenes(object = seurat_campbell_fasted,
                          mean.function = ExpMean,
                          dispersion.function = LogVMR,
                          x.low.cutoff = 0.02,
                          x.high.cutoff = 3,
                          y.cutoff = 0.5,
                          num.bin = 20,
                          binning.method = "equal_width"
                          )
                          
                          == 3006 variable genes


seurat_campbell_fasted <- FindVariableGenes(object = seurat_campbell_fasted,
                          mean.function = ExpMean,
                          dispersion.function = LogVMR,
                          x.low.cutoff = 0.02,
                          x.high.cutoff = 3,
                          y.cutoff = 0.5,
                          num.bin = 40,
                          binning.method = "equal_width"
                          )
                          
                          == 2870 variable gene

seurat_campbell_fasted <- FindVariableGenes(object = seurat_campbell_fasted,
                          mean.function = ExpMean,
                          dispersion.function = LogVMR,
                          x.low.cutoff = 0.02,
                          x.high.cutoff = 3,
                          y.cutoff = 0.5,
                          num.bin = 20,
                          binning.method = "equal_width"
                          )
                          
                          == 3006 variable genes

## Scale normalised gene expression per cell to remove unwanted sources of variation

Scale gene expression per cell by building linear models for nUMI, percent_mito and mouse replicate.

```{r ScaleData}

## Scale data nUMI, percent_mito and mouse replicate
seurat_campbell_fasted <- ScaleData(object = seurat_campbell_fasted, vars.to.regress = c("nUMI", "percent_mito", "replicate_name"))


```


### Principal component anlysis of variable genes.

Principal component anlysis of variable genes for use in cell clustering.

```{r RunPCA}

## Perform principal component analysis on variable genes
seurat_campbell_fasted <- RunPCA(object = seurat_campbell_fasted,
               pc.genes = seurat_campbell_fasted@var.genes,
               do.print = TRUE,
               pcs.print = 1:5,
               genes.print = 5)


## visualise top genes associated with principal components
VizPCA(object = seurat_campbell_fasted, pcs.use = 1:9)


## Plot principal component 1 v's 2
PCAPlot(object = seurat_campbell_fasted, dim.1 = 1, dim.2 = 2)


## Plot principal component 2 v's 3
PCAPlot(object = seurat_campbell_fasted, dim.1 = 2, dim.2 = 3)


## Plot heat map for gene expression of principal component 1 genes
PCHeatmap(object = seurat_campbell_fasted, pc.use = 1, cells.use = 500, do.balanced = TRUE, label.columns = FALSE)
# Ignore warnings!


## Plot heat map for gene expression of principal component 2 genes
PCHeatmap(object = seurat_campbell_fasted, pc.use = 2, cells.use = 500, do.balanced = TRUE, label.columns = FALSE)
# Ignore warnings!


## Plot heat map for gene expression of principal component 3 genes
PCHeatmap(object = seurat_campbell_fasted, pc.use = 3, cells.use = 500, do.balanced = TRUE, label.columns = FALSE)
# Ignore warnings!


## Plot heat map for gene expression of principal component 4 genes
PCHeatmap(object = seurat_campbell_fasted, pc.use = 4, cells.use = 500, do.balanced = TRUE, label.columns = FALSE)
# Ignore warnings!


## Plot multiple heatmaps of gene expression per PC genes
PCHeatmap(object = seurat_campbell_fasted,
          pc.use = 5:18,
          cells.use = 500,
          do.balanced = TRUE,
          label.columns = FALSE)


## Takes a pre-computed PCA (from most variable genes identified earlier) and projects this onto the entire dataset (all genes)
seurat_campbell_fasted <- ProjectPCA(object = seurat_campbell_fasted, do.print = TRUE)

```



### Determine statistically significant principal components

```{r}

## Perform jackstraw statistical test to investigate statistically significant PC.
seurat_campbell_fasted <- JackStraw(object = seurat_campbell_fasted,
                  num.replicate = 100,
                  display.progress = TRUE
                  )
# Maximum number of PCs allowed = 20.


## Visualise JackStraw plots
JackStrawPlot(object = seurat_campbell_fasted, PCs = 1:20)


## A less computationally intensive heuristic method for finding the statistically significant PCAs is using an elbow plot 
PCElbowPlot(object = seurat_campbell_fasted)


```



### Cell clustering

```{r}

## Cluster cells by PC
seurat_campbell_fasted <- FindClusters(object = seurat_campbell_fasted, reduction.type = "pca", dims.use = 1:20, resolution = 0.6, print.output = 0, save.SNN = TRUE, force.recalc = TRUE)

PrintFindClustersParams(object = seurat_campbell_fasted)

seurat_campbell_fasted <- RunTSNE(object = seurat_campbell_fasted, dims.use = 1:20, do.fast = TRUE)

# TSNEPlot(object = seurat_campbell_fasted)
TSNEPlot(object = seurat_campbell_fasted,
         no.legend = TRUE,
         do.label = TRUE)
## 1379 variable genes = 16 clusters
## 1379 variable genes, 10 PC = 13 clusters
## 3006 variable genes, 20 PC = 16 clusters

table(seurat_campbell_fasted@ident)

proportion_table<- table(seurat_campbell_fasted@ident, seurat_campbell_fasted@meta.data$replicate_name)

proportion_table<- round(prop.table(proportion_table, 2), 5)

proportion_table

TSNEPlot(object = seurat_campbell_fasted,
         group.by = "replicate_name",
         no.legend = FALSE,
         do.label = FALSE,
         pt.size = 0.4
         )

proportion_table<- data.frame(matrix(proportion_table, ncol = ncol(proportion_table)))

colnames(proportion_table)<- colnames(table(seurat_campbell_fasted@ident,
                                            seurat_campbell_fasted@meta.data$replicate_name))

chart.Correlation(data.frame( proportion_table[1:5,]) )

#######
seurat_campbell_fasted <- FindClusters(object = seurat_campbell_fasted, reduction.type = "pca", dims.use = 1:20, 
    resolution = 1.0, print.output = 0, save.SNN = TRUE, force.recalc = TRUE)

PrintFindClustersParams(object = seurat_campbell_fasted)

seurat_campbell_fasted <- RunTSNE(object = seurat_campbell_fasted, dims.use = 1:20, do.fast = TRUE)

# TSNEPlot(object = seurat_campbell_fasted)
TSNEPlot(object = seurat_campbell_fasted,
         no.legend = TRUE,
         do.label = TRUE)

##  1379 variable genes = 19 clusters
## 1379 variable genes, 10 PC = 15 clusters
## 3006 variable genes, 20 PC = 18 clusters

table(seurat_campbell_fasted@ident)

proportion_table<- table(seurat_campbell_fasted@ident, seurat_campbell_fasted@meta.data$replicate_name)

proportion_table<- round(prop.table(proportion_table, 2), 5)

proportion_table

TSNEPlot(object = seurat_campbell_fasted,
         group.by = "replicate_name",
         no.legend = FALSE,
         do.label = FALSE,
         pt.size = 0.4
         )


proportion_table<- data.frame(matrix(proportion_table, ncol = ncol(proportion_table)))

colnames(proportion_table)<- colnames(table(seurat_campbell_fasted@ident, 
                                            seurat_campbell_fasted@meta.data$replicate_name))

chart.Correlation(data.frame( proportion_table[1:5,]) )


#######
seurat_campbell_fasted <- FindClusters(object = seurat_campbell_fasted, reduction.type = "pca", dims.use = 1:20, 
    resolution = 1.5, print.output = 0, save.SNN = TRUE, force.recalc = TRUE)

PrintFindClustersParams(object = seurat_campbell_fasted)

seurat_campbell_fasted <- RunTSNE(object = seurat_campbell_fasted, dims.use = 1:20, do.fast = TRUE)

# TSNEPlot(object = seurat_campbell_fasted)
TSNEPlot(object = seurat_campbell_fasted,
         no.legend = TRUE,
         do.label = TRUE)

##  1379 variable genes = 20 clusters
## 1379 variable genes, 10 PC = 19 clusters
## 3006 variable genes, 20 PC = 20 clusters

table(seurat_campbell_fasted@ident)

proportion_table<- table(seurat_campbell_fasted@ident, seurat_campbell_fasted@meta.data$replicate_name)

proportion_table<- round(prop.table(proportion_table, 2), 5)

proportion_table

TSNEPlot(object = seurat_campbell_fasted,
         group.by = "replicate_name",
         no.legend = FALSE,
         do.label = FALSE,
         pt.size = 0.4
         )


proportion_table<- data.frame(matrix(proportion_table, ncol = ncol(proportion_table)))

colnames(proportion_table)<- colnames(table(seurat_campbell_fasted@ident, 
                                            seurat_campbell_fasted@meta.data$replicate_name))

chart.Correlation(data.frame( proportion_table[1:5,]) )

#######
seurat_campbell_fasted <- FindClusters(object = seurat_campbell_fasted, reduction.type = "pca", dims.use = 1:20, 
    resolution = 2.0, print.output = 0, save.SNN = TRUE, force.recalc = TRUE)

PrintFindClustersParams(object = seurat_campbell_fasted)

seurat_campbell_fasted <- RunTSNE(object = seurat_campbell_fasted, dims.use = 1:20, do.fast = TRUE)

# TSNEPlot(object = seurat_campbell_fasted)
TSNEPlot(object = seurat_campbell_fasted,
         no.legend = TRUE,
         do.label = TRUE)

##  1379 variable genes = 23 clusters
## 1379 variable genes, 10 PC = 21 clusters
## 3006 variable genes, 20 PC = 21 clusters

table(seurat_campbell_fasted@ident)

proportion_table<- table(seurat_campbell_fasted@ident, seurat_campbell_fasted@meta.data$replicate_name)

proportion_table<- round(prop.table(proportion_table, 2), 5)

proportion_table

TSNEPlot(object = seurat_campbell_fasted,
         group.by = "replicate_name",
         no.legend = FALSE,
         do.label = FALSE,
         pt.size = 0.4 )

proportion_table<- data.frame(matrix(proportion_table, ncol = ncol(proportion_table)))

colnames(proportion_table)<- colnames(table(seurat_campbell_fasted@ident, 
                                            seurat_campbell_fasted@meta.data$replicate_name))

chart.Correlation(data.frame( proportion_table[1:5,]) )

# cor(proportion_table[1],proportion_table[4])

## Cluster cells using final parameters (1817 genes, 20 PC, resolution = 0.6)
seurat_campbell_fasted <- FindClusters(object = seurat_campbell_fasted, reduction.type = "pca", dims.use = 1:20, resolution = 0.6, print.output = 0, save.SNN = TRUE, force.recalc = TRUE)


## Produce t-SNE of final parameters:
seurat_campbell_fasted <- RunTSNE(object = seurat_campbell_fasted, dims.use = 1:20, do.fast = TRUE)

```
 Use 1379 variable genes and 20 principal components with a resolution of 0.6. This gives a total of 16 clusters.


### Finding differentially expressed genes between cell clusters (cluster biomarkers)

```{r}


## Find markers for every cluster compared to all remaining cells, report both positive and negative genes.
seurat_campbell_fasted_biomarkers <- FindAllMarkers(object = seurat_campbell_fasted, only.pos = FALSE, min.pct = 0.2)


## Get the top 10 biomarkers per cluster
top10_seurat_campbell_markers<- seurat_campbell_fasted_biomarkers %>% group_by(cluster) %>% top_n(10, avg_logFC)
top10_seurat_campbell_markers

# write.csv(as.data.frame(seurat_campbell_fasted_biomarkers), file = "seurat_campbell_fasted_biomarkers.csv", quote = FALSE)
# write.csv(as.data.frame(top10_seurat_campbell_markers), file = "top10_seurat_campbell_fasted_biomarkers.csv", quote = FALSE)




## Perform ROC DE test. This can take a long time.
seurat_campbell_fasted_biomarkers_ROC <- FindAllMarkers(object = seurat_campbell_fasted, only.pos = FALSE, min.pct = 0.2, test.use = "roc")

top10_seurat_campbell_markers_ROC<- seurat_campbell_fasted_biomarkers_ROC %>% group_by(cluster) %>% top_n(10, avg_logFC)
top10_seurat_campbell_markers_ROC

# write.csv(as.data.frame(seurat_campbell_fasted_biomarkers_ROC), file = "seurat_campbell_fasted_biomarkers_ROC.csv", quote = FALSE)
# write.csv(as.data.frame(top10_seurat_campbell_markers_ROC), file = "top10_seurat_campbell_fasted_biomarkers_ROC.csv", quote = FALSE)


## Plot heatmap of top 10 DE genes
DoHeatmap(object = seurat_campbell_fasted,
          genes.use = top10_seurat_campbell_markers$gene,
          slim.col.label = TRUE,
          remove.key = TRUE)

```



### Save seurat object of chow mice.

```{r}

## save seurat object as .rds 
# saveRDS(seurat_campbell_fasted, file = "./seurat_campbell_fasted_final.rds")

```




















# Analysis of Campbell scRNAseq data for mice on a LOW FAT DIET(1 replicate) using seurat


### Load LFD mouse data

```{r readRDS}

## Load seurat object
seurat_campbell_lfd<- readRDS(file = "./seurat_campbell_lfd_just_created.rds")

seurat_campbell_lfd


```

### Quality control on CHOW mouse data

```{r QC}

## Import list of mouse mitocondrially encoded genes:
mito_genes<- read.csv("mito_genes_table.csv", header = TRUE)


## Make list of mitocondrial gene IDS:
mito_genes <- mito_genes$id_with_url


## Identify which mitochondrially expressed gene IDs are present in my dataset:
mito_genes_present<-seurat_campbell_lfd@raw.data[mito_genes, ]

mito_genes_present<- mito_genes[!grepl("NA", rownames(mito_genes_present))]

seurat_campbell_lfd@raw.data[mito_genes_present, 1:5]

dim(seurat_campbell_lfd@raw.data[mito_genes_present, ])


## Calculate the percentage of mitcondrial gene counts per cell
percent_mito <- Matrix::colSums(seurat_campbell_lfd@raw.data[mito_genes_present, ])/Matrix::colSums(seurat_campbell_lfd@raw.data)


## Basic stats of proportion of mitocondrial gene expression per cell
summary(percent_mito)


## add percentage mitocondrial genes into metadata
seurat_campbell_lfd <- AddMetaData(object = seurat_campbell_lfd,
                    metadata = percent_mito,
                    col.name = "percent_mito")


## Look at the seurat object meta data
head(seurat_campbell_lfd@meta.data)


## QC plots of number of genes, UMIs, and % mitochondria
VlnPlot(object = seurat_campbell_lfd,
        features.plot = c("nGene", "nUMI", "percent_mito"),
        nCol = 3,
        x.lab.rot = TRUE,
        point.size.use = 0.2
        )


## QC plots to show the relationship between nUMIs and relative mitocondrial gene expression or number of genes.
par(mfrow = c(1, 2))
# GenePlot(object = seurat_campbell_lfd, gene1 = "nUMI", gene2 = "percent_mito", do.hover = TRUE, pch.use = 1)
# GenePlot(object = seurat_campbell_lfd, gene1 = "nUMI", gene2 = "nGene", do.hover = TRUE, pch.use = 1)
GenePlot(object = seurat_campbell_lfd, gene1 = "nUMI", gene2 = "percent_mito", pch.use = 16, cex.use = 0.5)

GenePlot(object = seurat_campbell_lfd, gene1 = "nUMI", gene2 = "nGene", pch.use = 16, cex.use = 0.5)


```

### Filter cells after QC

Filter cells out with more than 0.4% of total gene expression comming from mitocondrially encoded genes and more than 5000 genes expressed. 

```{r FilterCells}

## manual check; I already know all cells have >800 genes
table(seurat_campbell_lfd@meta.data$percent_mito < 0.004 & seurat_campbell_lfd@meta.data$nGene<5000)
 

# FALSE  TRUE 
#   213  3134 
 
## Filter cells with <0.4% percent_mito and <5000 genes
seurat_campbell_lfd <- FilterCells(object = seurat_campbell_lfd,
                    subset.names = c("nGene", "percent_mito"),
                    low.thresholds = c(800, -Inf),
                    high.thresholds = c(5000, 0.004))
 
seurat_campbell_lfd
# An object of class seurat in project CAMPBELL_FASTED 
# 21789 genes across 3640 samples.

 
```


### Log normalise gene expression per cell

```{r NormalizeData}

## Plot graph of total expression before normalisation
hist(colSums(seurat_campbell_lfd@data),
     breaks = 100,
     main = "Total expression before normalisation",
     xlab = "Sum of expression")

## Normalise gene expression per cell
seurat_campbell_lfd <- NormalizeData(object = seurat_campbell_lfd, normalization.method = "LogNormalize", scale.factor = 10000)

## Plot graph of total expression after normalisation
hist(colSums(as.data.frame(as.matrix(seurat_campbell_lfd@data))),
     breaks = 100,
     main = "Total expression after normalisation",
     xlab = "Sum of expression")


```



### Find genes whose expression varies between cells.

Find genes whose expression varies between cells, which will be used to construct principal componets between cells that will be used for clustering.

```{r FindVariableGenes}

## Find variable genes by expression
seurat_campbell_lfd <- FindVariableGenes(object = seurat_campbell_lfd,
                          mean.function = ExpMean,
                          dispersion.function = LogVMR,
                          x.low.cutoff = 0.05,
                          x.high.cutoff = 3,
                          y.cutoff = 0.75,
                          num.bin = 20,
                          binning.method = "equal_width"
                          )

# number of variable genes
length(seurat_campbell_lfd@var.genes)

```

seurat_campbell_lfd <- FindVariableGenes(object = seurat_campbell_lfd,
                          mean.function = ExpMean,
                          dispersion.function = LogVMR,
                          x.low.cutoff = 0.05,
                          x.high.cutoff = 4,
                          y.cutoff = 0.75,
                          num.bin = 20,
                          binning.method = "equal_width"
                          )
                          
                          == 1484 variable genes


seurat_campbell_lfd <- FindVariableGenes(object = seurat_campbell_lfd,
                          mean.function = ExpMean,
                          dispersion.function = LogVMR,
                          x.low.cutoff = 0.05,
                          x.high.cutoff = 4,
                          y.cutoff = 0.5,
                          num.bin = 20,
                          binning.method = "equal_width"
                          )
                          
                          == 2420 variable genes

seurat_campbell_lfd <- FindVariableGenes(object = seurat_campbell_lfd,
                          mean.function = ExpMean,
                          dispersion.function = LogVMR,
                          x.low.cutoff = 0.02,
                          x.high.cutoff = 3,
                          y.cutoff = 0.75,
                          num.bin = 20,
                          binning.method = "equal_width"
                          )
                          
                          == 1986 variable genes

seurat_campbell_lfd <- FindVariableGenes(object = seurat_campbell_lfd,
                          mean.function = ExpMean,
                          dispersion.function = LogVMR,
                          x.low.cutoff = 0.02,
                          x.high.cutoff = 3,
                          y.cutoff = 0.5,
                          num.bin = 20,
                          binning.method = "equal_width"
                          )
                          
                          == 3189 variable genes


seurat_campbell_lfd <- FindVariableGenes(object = seurat_campbell_lfd,
                          mean.function = ExpMean,
                          dispersion.function = LogVMR,
                          x.low.cutoff = 0.02,
                          x.high.cutoff = 3,
                          y.cutoff = 0.5,
                          num.bin = 40,
                          binning.method = "equal_width"
                          )
                          
                          == 3122 variable gene


seurat_campbell_lfd <- FindVariableGenes(object = seurat_campbell_lfd,
                          mean.function = ExpMean,
                          dispersion.function = LogVMR,
                          x.low.cutoff = 0.05,
                          x.high.cutoff = 3,
                          y.cutoff = 0.75,
                          num.bin = 20,
                          binning.method = "equal_width"
                          )
                          
                          == 1480 variable gene
                          

## Scale normalised gene expression per cell to remove unwanted sources of variation

Scale gene expression per cell by building linear models for nUMI, percent_mito and mouse replicate.

```{r ScaleData}

## Scale data nUMI, percent_mito and mouse replicate
seurat_campbell_lfd <- ScaleData(object = seurat_campbell_lfd, vars.to.regress = c("nUMI", "percent_mito"))


```


### Principal component anlysis of variable genes.

Principal component anlysis of variable genes for use in cell clustering.

```{r RunPCA}

## Perform principal component analysis on variable genes
seurat_campbell_lfd <- RunPCA(object = seurat_campbell_lfd,
               pc.genes = seurat_campbell_lfd@var.genes,
               do.print = TRUE,
               pcs.print = 1:5,
               genes.print = 5)


## visualise top genes associated with principal components
VizPCA(object = seurat_campbell_lfd, pcs.use = 1:9)


## Plot principal component 1 v's 2
PCAPlot(object = seurat_campbell_lfd, dim.1 = 1, dim.2 = 2)


## Plot principal component 2 v's 3
PCAPlot(object = seurat_campbell_lfd, dim.1 = 2, dim.2 = 3)


## Plot heat map for gene expression of principal component 1 genes
PCHeatmap(object = seurat_campbell_lfd, pc.use = 1, cells.use = 500, do.balanced = TRUE, label.columns = FALSE)
# Ignore warnings!


## Plot heat map for gene expression of principal component 2 genes
PCHeatmap(object = seurat_campbell_lfd, pc.use = 2, cells.use = 500, do.balanced = TRUE, label.columns = FALSE)
# Ignore warnings!


## Plot heat map for gene expression of principal component 3 genes
PCHeatmap(object = seurat_campbell_lfd, pc.use = 3, cells.use = 500, do.balanced = TRUE, label.columns = FALSE)
# Ignore warnings!


## Plot heat map for gene expression of principal component 4 genes
PCHeatmap(object = seurat_campbell_lfd, pc.use = 4, cells.use = 500, do.balanced = TRUE, label.columns = FALSE)
# Ignore warnings!


## Plot multiple heatmaps of gene expression per PC genes
PCHeatmap(object = seurat_campbell_lfd,
          pc.use = 5:18,
          cells.use = 500,
          do.balanced = TRUE,
          label.columns = FALSE)


## Takes a pre-computed PCA (from most variable genes identified earlier) and projects this onto the entire dataset (all genes)
seurat_campbell_lfd <- ProjectPCA(object = seurat_campbell_lfd, do.print = TRUE)

```



### Determine statistically significant principal components

```{r}

## Perform jackstraw statistical test to investigate statistically significant PC.
seurat_campbell_lfd <- JackStraw(object = seurat_campbell_lfd,
                  num.replicate = 100,
                  display.progress = TRUE
                  )
# Maximum number of PCs allowed = 20.


## Visualise JackStraw plots
JackStrawPlot(object = seurat_campbell_lfd, PCs = 1:20)


## A less computationally intensive heuristic method for finding the statistically significant PCAs is using an elbow plot 
PCElbowPlot(object = seurat_campbell_lfd)


```



### Cell clustering

```{r}

## Cluster cells by PC
seurat_campbell_lfd <- FindClusters(object = seurat_campbell_lfd, reduction.type = "pca", dims.use = 1:16, resolution = 0.6, print.output = 0, save.SNN = TRUE, force.recalc = TRUE)

PrintFindClustersParams(object = seurat_campbell_lfd)

seurat_campbell_lfd <- RunTSNE(object = seurat_campbell_lfd, dims.use = 1:16, do.fast = TRUE)

# TSNEPlot(object = seurat_campbell_lfd)
TSNEPlot(object = seurat_campbell_lfd,
         no.legend = TRUE,
         do.label = TRUE)
## 1986 variable genes, 20 PC = 14 clusters
## 1986 variable genes, 16 PC = 14 clusters

table(seurat_campbell_lfd@ident)


#######
seurat_campbell_lfd <- FindClusters(object = seurat_campbell_lfd, reduction.type = "pca", dims.use = 1:16, 
    resolution = 1.0, print.output = 0, save.SNN = TRUE, force.recalc = TRUE)

PrintFindClustersParams(object = seurat_campbell_lfd)

seurat_campbell_lfd <- RunTSNE(object = seurat_campbell_lfd, dims.use = 1:16, do.fast = TRUE)

# TSNEPlot(object = seurat_campbell_lfd)
TSNEPlot(object = seurat_campbell_lfd,
         no.legend = TRUE,
         do.label = TRUE)

##  1986 variable genes, 20 PC = 18 clusters
## 1986 variable genes, 16 PC = 16 clusters

table(seurat_campbell_lfd@ident)


#######
seurat_campbell_lfd <- FindClusters(object = seurat_campbell_lfd, reduction.type = "pca", dims.use = 1:16, 
    resolution = 1.5, print.output = 0, save.SNN = TRUE, force.recalc = TRUE)

PrintFindClustersParams(object = seurat_campbell_lfd)

seurat_campbell_lfd <- RunTSNE(object = seurat_campbell_lfd, dims.use = 1:16, do.fast = TRUE)

# TSNEPlot(object = seurat_campbell_lfd)
TSNEPlot(object = seurat_campbell_lfd,
         no.legend = TRUE,
         do.label = TRUE)

##  1986 variable genes, 20 PC = 18 clusters
## 1986 variable genes, 16 PC = 18 clusters

table(seurat_campbell_lfd@ident)


#######
seurat_campbell_lfd <- FindClusters(object = seurat_campbell_lfd, reduction.type = "pca", dims.use = 1:16, 
    resolution = 2.0, print.output = 0, save.SNN = TRUE, force.recalc = TRUE)

PrintFindClustersParams(object = seurat_campbell_lfd)

seurat_campbell_lfd <- RunTSNE(object = seurat_campbell_lfd, dims.use = 1:16, do.fast = TRUE)

# TSNEPlot(object = seurat_campbell_lfd)
TSNEPlot(object = seurat_campbell_lfd,
         no.legend = TRUE,
         do.label = TRUE)

##  1986 variable genes, 20 PC = 20 clusters
## 1986 variable genes, 16 PC = 20 clusters

table(seurat_campbell_lfd@ident)


## Cluster cells using final parameters (1986 genes, 16 PC, resolution = 1.0)
seurat_campbell_lfd <- FindClusters(object = seurat_campbell_lfd, reduction.type = "pca", dims.use = 1:16, resolution = 1.0, print.output = 0, save.SNN = TRUE, force.recalc = TRUE)


## Produce t-SNE of final parameters:
seurat_campbell_lfd <- RunTSNE(object = seurat_campbell_lfd, dims.use = 1:16, do.fast = TRUE)

```
 Use 1986 variable genes and 16 principal components with a resolution of 1.0 (Changing number of PC makes very little difference to the clustering). This gives a total of 16 clusters.


### Finding differentially expressed genes between cell clusters (cluster biomarkers)

```{r}


## Find markers for every cluster compared to all remaining cells, report both positive and negative genes.
seurat_campbell_lfd_biomarkers <- FindAllMarkers(object = seurat_campbell_lfd, only.pos = FALSE, min.pct = 0.2)


## Get the top 10 biomarkers per cluster
top10_seurat_campbell_markers<- seurat_campbell_lfd_biomarkers %>% group_by(cluster) %>% top_n(10, avg_logFC)
top10_seurat_campbell_markers

write.csv(as.data.frame(seurat_campbell_lfd_biomarkers), file = "seurat_campbell_lfd_biomarkers.csv", quote = FALSE)
write.csv(as.data.frame(top10_seurat_campbell_markers), file = "top10_seurat_campbell_lfd_biomarkers.csv", quote = FALSE)




## Perform ROC DE test. This can take a long time.
seurat_campbell_lfd_biomarkers_ROC <- FindAllMarkers(object = seurat_campbell_lfd, only.pos = FALSE, min.pct = 0.2, test.use = "roc")

top10_seurat_campbell_markers_ROC<- seurat_campbell_lfd_biomarkers_ROC %>% group_by(cluster) %>% top_n(10, avg_logFC)
top10_seurat_campbell_markers_ROC

write.csv(as.data.frame(seurat_campbell_lfd_biomarkers_ROC), file = "seurat_campbell_lfd_biomarkers_ROC.csv", quote = FALSE)
write.csv(as.data.frame(top10_seurat_campbell_markers_ROC), file = "top10_seurat_campbell_lfd_biomarkers_ROC.csv", quote = FALSE)


## Plot heatmap of top 10 DE genes
DoHeatmap(object = seurat_campbell_lfd,
          genes.use = top10_seurat_campbell_markers$gene,
          slim.col.label = TRUE,
          remove.key = TRUE)

```



### Save seurat object of chow mice.

```{r}

## save seurat object as .rds 
saveRDS(seurat_campbell_lfd, file = "./seurat_campbell_lfd_final.rds")

```















### title

```{r}

## 


```



### title

```{r}

## 


```



### title

```{r}

## 


```



### title

```{r}

## 


```



### title

```{r}

## 


```



### title

```{r}

## 


```



### title

```{r}

## 


```



### title

```{r}

## 


```




### title

```{r}

## 


```



### title

```{r}

## 


```


### title

```{r}

## 


```


### title

```{r}

## 


```


### title

```{r}

## 


```
