---
title: "Campbell_2017_prepare_expression_matrices"
author: "Matthew Banton"
date: "14/11/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load packages and set working directory

```{r}

library(tidyverse)
library(scater)
library(dplyr)
library(limma)
library(SingleCellExperiment)
library(Seurat)
library(mclust)
library(Matrix)
library(tibble)

setwd("/media/data/mattb/projects/Brian_scRNAseq_website/R_projects/mouse_hypothalamus_scRNAseq")

```



## Create function to import and join campbell expression matrices created by zUMIs

From the cambell paper the data for the following mouse replicates were produced:

* CHOW fed: SRR5164436, SRR5164437, SRR5164438, SRR5164443, SRR5164445
* Low fat diet: SRR5164439
* High fat diet: SRR5164440
* Fasted: SRR5164441, SRR5164444, SRR5164446
* Refed: SRR5164442

A function has been created that can import and join expression matrices from different mice treated with the same conditions.

```{r}

## for loop function that loads files and joins them together into one seurat object:

Create_joined_seurat_object<- function(file_list) {



  ## Process the first file:
 file_to_process<- file_list[1]

  ## Load object
  campbell_zUMIs_object<- readRDS(paste0("/media/data/mattb/projects/Brian_scRNAseq_website/data/campbell/", file_to_process, "/zUMIs_output/expression/", file_to_process, "_default_STAR.dgecounts.rds"))
  
  ## extract counts from object
  first_table<-  as.data.frame(as.matrix(campbell_zUMIs_object$umicount$exon$all)) 
  
  ## Create new column names
 colnames(first_table)<- paste0(file_to_process, "_",                                                          colnames(first_table))
  
 ## Use tibble function to move rownames to a column
 first_table<- rownames_to_column(first_table, var = "gene_ID")


## For loop to process the rest of the files:
for (i in 2:length(file_list)) {
  
  ## Process the first file:
 file_to_process<- file_list[i]

  ## Load object
  campbell_zUMIs_object<- readRDS(paste0("/media/data/mattb/projects/Brian_scRNAseq_website/data/campbell/", file_to_process, "/zUMIs_output/expression/", file_to_process, "_default_STAR.dgecounts.rds"))
  
  ## extract counts from object
  second_table<-  as.data.frame(as.matrix(campbell_zUMIs_object$umicount$exon$all)) 
  
  ## Create new column names
 colnames(second_table)<- paste0(file_to_process, "_",                                                          colnames(second_table))
  
 ## Use tibble function to move rownames to a column
 second_table<- rownames_to_column(second_table, var = "gene_ID")

  ## join tables together
  first_table<- dplyr::full_join(first_table, second_table, by = "gene_ID")
  
}

final_table<- column_to_rownames(first_table, var = "gene_ID" )

return(final_table)

}


```


## Create and save a seurat object contaiing all five replicates for mice fed on normal CHOW.

```{r}


## Create CHOW expression matrix:
CHOW_file_list<- c("SRR5164436", "SRR5164437", "SRR5164438", "SRR5164443", "SRR5164445")


## Use function to join expression matrices
campbell_CHOW_expression_matrix<- Create_joined_seurat_object(CHOW_file_list)

dim(campbell_CHOW_expression_matrix)

campbell_CHOW_expression_matrix[c(1:6, 19000:19005, 24000:24005),c(1:6, 19000:19005, 24000:24005)]


###########
## Replace na's in table to 0:

campbell_CHOW_expression_matrix[is.na(campbell_CHOW_expression_matrix)] <- 0 # Despite what the forums say this method was quickist

dim(campbell_CHOW_expression_matrix)

campbell_CHOW_expression_matrix[c(1:6, 19000:19005, 24000:24005),c(1:6, 19000:19005, 24000:24005)]


#############
## Histogram of how many genes are expressed in each cell : 
at_least_one <- apply(campbell_CHOW_expression_matrix, 2, function(x) sum(x>0))

head(at_least_one)

par(mfrow=c(1,2)) # One row and two columns

hist(at_least_one, breaks = 100,
     main = "Distribution of detected genes",
     xlab = "Genes with at least one tag")
abline(v= 800, col = "red") 


## Histogram of number of counts per cell: 
colsums_vector<-colSums(campbell_CHOW_expression_matrix)

head(colsums_vector)

hist(colsums_vector,
     breaks = 100, main = "Expression sum per cell",
     xlab = "Sum expression")


############
## Create a Seurat object to contain scRNA seq data:
seurat_campbell_chow <- CreateSeuratObject(raw.data = campbell_CHOW_expression_matrix, 
                                 min.cells = 3, 
                                 min.genes = 800, 
                                 project = "CAMPBELL_CHOW")


# How many cells and genes passed through initial filtering:
seurat_campbell_chow
# from : 35573 cells and 37406 genes
# to : 11255 cells and 30000 genes


##############
## Add mouse replicate info into  metadata

## Investigate the meta data
head(seurat_campbell_chow@meta.data)


## Create vector with mouse replicate name i.e. name of SRR... fastq file 
replicate_name<- rownames(seurat_campbell_chow@meta.data)

names(replicate_name)<- replicate_name

replicate_name<-gsub("\\_.*", "", replicate_name)

## Add metadata to seurat object
seurat_campbell_chow <- AddMetaData(object = seurat_campbell_chow,
                    metadata = replicate_name,
                    col.name = "replicate_name")


# Investigate the meta data
head(seurat_campbell_chow@meta.data)
seurat_campbell_chow@meta.data[c(1:5, 11001:11005),]


## save object:
# saveRDS(seurat_campbell_chow, file = "./seurat_campbell_chow_just_created.rds")



```


## Create and save a seurat object containing all three replicates for fasted mice.


```{r}


## Create fasted expression matrix:
fasted_file_list<- c("SRR5164441", "SRR5164444", "SRR5164446")


## Use function to join expression matrices
campbell_fasted_expression_matrix<- Create_joined_seurat_object(fasted_file_list)

dim(campbell_fasted_expression_matrix)
# [1] 31887 35277

campbell_fasted_expression_matrix[c(1:6, 19000:19005, 24000:24005),c(1:6, 19000:19005, 24000:24005)]


###########
## Replace na's in table to 0:

campbell_fasted_expression_matrix[is.na(campbell_fasted_expression_matrix)] <- 0 # Despite what the forums say this method was quickist

dim(campbell_fasted_expression_matrix)

campbell_fasted_expression_matrix[c(1:6, 19000:19005, 24000:24005),c(1:6, 19000:19005, 24000:24005)]


#############
## Hsitogram of how many genes are expressed in each cell : 
at_least_one <- apply(campbell_fasted_expression_matrix, 2, function(x) sum(x>0))

summary (at_least_one)
#   Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
#    1.0   245.0   325.0   454.9   499.0  7195.0 
    
head(at_least_one)

par(mfrow=c(1,2)) # One row and two columns

hist(at_least_one, breaks = 100,
     main = "Distribution of detected genes",
     xlab = "Genes with at least one tag")
abline(v= 800, col = "red") 


## Histogram of number of counts per cell: 
colsums_vector<-colSums(campbell_fasted_expression_matrix)

summary(colsums_vector)
#    Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
#      2     312     421     676     676   30958 

head(colsums_vector)

hist(colsums_vector,
     breaks = 100, main = "Expression sum per cell",
     xlab = "Sum expression")


############
## Create a Seurat object to contain scRNA seq data:
seurat_campbell_fasted <- CreateSeuratObject(raw.data = campbell_fasted_expression_matrix, 
                                 min.cells = 3, 
                                 min.genes = 800, 
                                 project = "CAMPBELL_FASTED")


# How many cells and genes passed through initial filtering:
seurat_campbell_fasted
# from : 35277 cells and 31887 genes
# to : 3783 cells and 21789 genes


##############
## Add mouse replicate info into  metadata

## Investigate the meta data
head(seurat_campbell_fasted@meta.data)


## Create vector with mouse replicate name i.e. name of SRR... fastq file 
replicate_name<- rownames(seurat_campbell_fasted@meta.data)

names(replicate_name)<- replicate_name

replicate_name<-gsub("\\_.*", "", replicate_name)

## Add metadata to seurat object
seurat_campbell_fasted <- AddMetaData(object = seurat_campbell_fasted,
                    metadata = replicate_name,
                    col.name = "replicate_name")


# Investigate the meta data
head(seurat_campbell_fasted@meta.data)
seurat_campbell_fasted@meta.data[c(1:5, 3001:3005),]


## save object:
# saveRDS(seurat_campbell_fasted, file = "./seurat_campbell_fasted_just_created.rds")




```


## Create and save a seurat object containing a single replicate of a mouse on a High fat diet.


```{r}
# High fat diet: SRR5164440

## Load single R object from zUMIs containing the single expression matrix
file_to_process<- "SRR5164440"

campbell_zUMIs_object<- readRDS(paste0("/media/data/mattb/projects/Brian_scRNAseq_website/data/campbell/", file_to_process, "/zUMIs_output/expression/", file_to_process, "_default_STAR.dgecounts.rds"))
  
  ## extract counts from object
  campbell_hfd_expression_matrix<-  as.data.frame(as.matrix(campbell_zUMIs_object$umicount$exon$all)) 
  
dim(campbell_hfd_expression_matrix)
# 30504  6329

campbell_hfd_expression_matrix[1:6, 1:6]


#############
## Histogram of how many genes are expressed in each cell : 
at_least_one <- apply(campbell_hfd_expression_matrix, 2, function(x) sum(x>0))

summary (at_least_one)
#     Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
#      1     449    1061    1459    2097    9514 
    
head(at_least_one)

par(mfrow=c(1,2)) # One row and two columns

hist(at_least_one, breaks = 100,
     main = "Distribution of detected genes",
     xlab = "Genes with at least one tag")
abline(v= 800, col = "red") 


## Histogram of number of counts per cell: 
colsums_vector<-colSums(campbell_hfd_expression_matrix)

summary(colsums_vector)
#    Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
#     1     614    1617    2993    3848   57289 

head(colsums_vector)

hist(colsums_vector,
     breaks = 100, main = "Expression sum per cell",
     xlab = "Sum expression")


############
## Create a Seurat object to contain scRNA seq data:
seurat_campbell_hfd <- CreateSeuratObject(raw.data = campbell_hfd_expression_matrix, 
                                 min.cells = 3, 
                                 min.genes = 800, 
                                 project = "CAMPBELL_HFD")


# How many cells and genes passed through initial filtering:
seurat_campbell_hfd
# from : 6329 cells and 30504 genes
# to : 3778 cells and 23685 genes


##############
## Add mouse replicate info into  metadata

## Investigate the meta data
head(seurat_campbell_hfd@meta.data)


## Create vector with mouse replicate name i.e. name of SRR... fastq file 
replicate_name<- rep(file_to_process, nrow(seurat_campbell_hfd@meta.data))

names(replicate_name)<- rownames(seurat_campbell_hfd@meta.data)

seurat_campbell_hfd <- AddMetaData(object = seurat_campbell_hfd,
                    metadata = replicate_name,
                    col.name = "replicate_name")

# Investigate the meta data
head(seurat_campbell_hfd@meta.data)

seurat_campbell_hfd@meta.data[c(1:5, 3001:3005),]


## save object:
# saveRDS(seurat_campbell_hfd, file = "./seurat_campbell_hfd_just_created.rds")




```


## Create and save a seurat object containing a single replicate of a mouse on a Low fat diet.


```{r}
# Low fat diet: SRR5164439

## Load single R object from zUMIs containing the single expression matrix
file_to_process<- "SRR5164439"

campbell_zUMIs_object<- readRDS(paste0("/media/data/mattb/projects/Brian_scRNAseq_website/data/campbell/", file_to_process, "/zUMIs_output/expression/", file_to_process, "_default_STAR.dgecounts.rds"))
  
## extract counts from object
campbell_lfd_expression_matrix<-  as.data.frame(as.matrix(campbell_zUMIs_object$umicount$exon$all)) 
  
dim(campbell_lfd_expression_matrix)
# 29988  5353

campbell_lfd_expression_matrix[1:6, 1:6]


#############
## Histogram of how many genes are expressed in each cell : 
at_least_one <- apply(campbell_lfd_expression_matrix, 2, function(x) sum(x>0))

summary (at_least_one)
#   Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
#     3     533    1130    1506    2071    8471 
    
head(at_least_one)

par(mfrow=c(1,2)) # One row and two columns

hist(at_least_one, breaks = 100,
     main = "Distribution of detected genes",
     xlab = "Genes with at least one tag")
abline(v= 800, col = "red") 


## Histogram of number of counts per cell: 
colsums_vector<-colSums(campbell_lfd_expression_matrix)

summary(colsums_vector)
#   Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
#     7     724    1721    3109    3810   38435 

head(colsums_vector)

hist(colsums_vector,
     breaks = 100, main = "Expression sum per cell",
     xlab = "Sum expression")


############
## Create a Seurat object to contain scRNA seq data:
seurat_campbell_lfd <- CreateSeuratObject(raw.data = campbell_lfd_expression_matrix, 
                                 min.cells = 3, 
                                 min.genes = 800, 
                                 project = "CAMPBELL_LFD")


# How many cells and genes passed through initial filtering:
seurat_campbell_lfd
# from : 5353 cells and 29988 genes
# to : 3347 cells and 23273 genes

  
##############
## Add mouse replicate info into  metadata

## Investigate the meta data
head(seurat_campbell_lfd@meta.data)


## Create vector with mouse replicate name i.e. name of SRR... fastq file 
replicate_name<- rep(file_to_process, nrow(seurat_campbell_lfd@meta.data))

names(replicate_name)<- rownames(seurat_campbell_lfd@meta.data)

seurat_campbell_lfd <- AddMetaData(object = seurat_campbell_lfd,
                    metadata = replicate_name,
                    col.name = "replicate_name")

# Investigate the meta data
head(seurat_campbell_lfd@meta.data)

seurat_campbell_lfd@meta.data[c(1:5, 3001:3005),]


## save object:
# saveRDS(seurat_campbell_lfd, file = "./seurat_campbell_lfd_just_created.rds")



```


## Create and save a seurat object containing a single replicate of a mouse on a fasted diet that was subsequenctly refed.



```{r}
# Refed: SRR5164442

## Load single R object from zUMIs containing the single expression matrix
file_to_process<- "SRR5164442"

campbell_zUMIs_object<- readRDS(paste0("/media/data/mattb/projects/Brian_scRNAseq_website/data/campbell/", file_to_process, "/zUMIs_output/expression/", file_to_process, "_default_STAR.dgecounts.rds"))
  
  ## extract counts from object
  campbell_refed_expression_matrix<-  as.data.frame(as.matrix(campbell_zUMIs_object$umicount$exon$all)) 
  
dim(campbell_refed_expression_matrix)
#  28295 17928

campbell_refed_expression_matrix[1:6, 1:6]


#############
## Histogram of how many genes are expressed in each cell : 
at_least_one <- apply(campbell_refed_expression_matrix, 2, function(x) sum(x>0))

summary (at_least_one)
#   Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
#    1.0   298.0   388.0   517.9   579.0  6018.0  
    
head(at_least_one)

par(mfrow=c(1,2)) # One row and two columns

hist(at_least_one, breaks = 100,
     main = "Distribution of detected genes",
     xlab = "Genes with at least one tag")
abline(v= 800, col = "red") 


## Histogram of number of counts per cell: 
colsums_vector<-colSums(campbell_refed_expression_matrix)

summary(colsums_vector)
#    Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
#    5.0   373.0   495.0   744.7   778.0 17671.0 

head(colsums_vector)

hist(colsums_vector,
     breaks = 100, main = "Expression sum per cell",
     xlab = "Sum expression")


############
## Create a Seurat object to contain scRNA seq data:
seurat_campbell_refed <- CreateSeuratObject(raw.data = campbell_refed_expression_matrix, 
                                 min.cells = 3, 
                                 min.genes = 800, 
                                 project = "CAMPBELL_REFED")


# How many cells and genes passed through initial filtering:
seurat_campbell_refed
# from : 17928 cells and 28295 genes
# to : 2375 cells and 19267 genes



##############
## Add mouse replicate info into  metadata

## Investigate the meta data
head(seurat_campbell_refed@meta.data)


## Create vector with mouse replicate name i.e. name of SRR... fastq file 
replicate_name<- rep(file_to_process, nrow(seurat_campbell_refed@meta.data))

names(replicate_name)<- rownames(seurat_campbell_refed@meta.data)

seurat_campbell_refed <- AddMetaData(object = seurat_campbell_refed,
                    metadata = replicate_name,
                    col.name = "replicate_name")

# Investigate the meta data
head(seurat_campbell_refed@meta.data)

seurat_campbell_refed@meta.data[c(1:5, 2001:2005),]


## save object:
# saveRDS(seurat_campbell_refed, file = "./seurat_campbell_refed_just_created.rds")






```



